Phylogenetics and the human microbiome.

### 2013-09-24 sbreview.tex
More rearrangin'

more virome
http://www.pnas.org/content/early/2013/07/03/1300833110.short
review paper:http://genomebiology.com/2013/14/7/127

mycobiome
http://genomemedicine.com/content/5/7/63

@@

So far we have been focusing on bacteria, though archaea also play a role.
Archaea have also been found in the gut.
and fungi too
http://www.plosone.org/article/info%3Adoi%2F10.1371%2Fjournal.pone.0066019

Although the focus of much of the work has been on bacteria and archaea, other groups of life have attracted more attention lately, such as the fungi.

Viruses too
febrile study \cite{wylie2012sequence}

@@

\subsection{Data generated from the various studies}
reference genomes
HMP DACC.

@@

@@

\subsection{Software}
\cite{caporaso2010qiime}

\cite{schloss2009introducing}
Introducing mothur: open-source, platform-independen

\cite{lynch2013axiome}
AXIOME: automated exploration of microbial diversity

% databases
\cite{chen2010human}
The Human Oral Microbiome Database: a web accessible resource for investigating oral microbe taxonomic and genomic information

@@

But in addition to deeper probing we can also expect wider coverage.
Lots of data.
Citizen microbiome projects.

@@

Microbial outbreak tracking.
Very common in RNA viruses.
Already using similar methods by applying whole-genome sequencing.

>>>>

@@

We will be focusing on bacterial, as they have formed the bulk of the research.
However, archaea \cite{hoffmann2013archaea} and fungi play a role.
Viruses too \citep{reyes2010viruses,minot2011human,wylie2012sequence,minot2013rapid}.

@@

\subsection{Whole genomes}
Whole-genome sequencing used for microbial outbreak tracking \cite{snitkin2012tracking}.
Now from culture, but next we will see single-cell sequencing.

\subsection{Data availability and future}
reference genomes
HMP DACC.

@@

Future will give deeper sequencing, which can mean massive parallelization.
Citizen microbiome projects.

@@

\subsection{Software}

People do most phylogenetic inferences as part of a pipeline.
The most popular ones are QIIME \cite{caporaso2010qiime} and mothur \cite{schloss2009introducing}.
AXIOME streamlines and manages analysis of small subunit (SSU) rRNA marker data in QIIME and mothur \cite{lynch2013axiome}.

mothur has ported in the clearcut \cite{evans2006relaxed,sheneman2006clearcut} program.
QIIME wraps clearcut and \cite{price2010fasttree}.

Bayesian inference is very rare.

Databases:
\cite{chen2010human}
The Human Oral Microbiome Database: a web accessible resource for investigating oral microbe taxonomic and genomic information

@@

@@
